---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tricks

RStudio addins are handy but have flaws :

* They're contained in packages, which are not straightforward to write/iterate on
* It's hard to remember which addins you've installed, and harder to remember the
hotkeys you've set for them
* The addin list soon becomes overwhelming since they're all shown at all times,
even when they don't apply

So in the end you might not write them much, and not use them much.

{tricks} is an attempt to solve those issues:

* It works with a single hotkey
* Depending on your selection (or current line or document), only relevant action choices are proposed
* It's made very easy to add actions

We don't recommend attaching the package, a call to `tricks::add_tricks()` in your R
profile is all you should need.

## Installation

```{r, , eval = FALSE}
remotes::install_github("moodymudskipper/tricks")
```

## general

We call "tricks" the features implemented by {tricks}, and use the following syntax
to define them :

```{r, , eval = FALSE}
tricks::add_tricks(
  "<label1>" = <condition1> ~ <action1>, 
  "<label2>" = <condition2> ~ <action2>, 
  ...)
```

* When the **hotkey** is triggered , all **conditions** are evaluated
* When the **condition** is satisfied, the **label** of the trick is is shown
* When the **label** is selected, the **action** is triggered

The package contains a suit of functions to help you define actions and conditions conveniently.

Let's go through a few examples.

## Editing your R Profile

Do you know how to easily edit your user or project R Profile ? There's a nice {usethis} function
for this, but maybe you have problems remembering it, or you'd rather not have to
type it.

We propose below tricks to do that. If you trigger {tricks} when
not selecting anything in the editor it will propose you to edit either your user
or project R profile.

```{r, eval = FALSE}
tricks::add_tricks(
  "Edit user '.Rprofile'" =
    selection_is_empty() ~     # condition : no selection
    usethis::edit_r_profile(), # action    : call packaged function

  "Edit project '.Rprofile'" =
    selection_is_empty() ~                      # condition : no selection
    usethis::edit_r_profile(scope = "project"), # action    : call packaged function
)
```

We see below that whenever the selection is not empty the 
labels are not shown, but they are shown if the selection is empty.

*INSERT GIF*

Setting up tricks when the selection is empty is a good way to keep a list of
bookmarks for actions you don't want to remember how to trigger, maybe some other
{usethis} workflow actions.

`selection_is_empty()` is called a **condition helper**, and we have many of them,
documented in : 

* `?``selection-condition-helpers```
* `?``file-condition-helpers```
* `?``project-condition-helpers```
* `?``clipboard-condition-helpers```
* `?``system-condition-helpers```

## Reprex your selection

The {reprex} package comes with a handy addin to create a reprex from a hotkey.
You'll have to remember how to trigger the hotkey though.

If you don't want to we can set up a trick that will be proposed whenever we select
parsable code that isn't a symbol, when it is triggered it should call the relevant {reprex} addin.

```{r, eval = FALSE}
tricks::add_tricks(
  "Reprex selection" =
    selection_is_parsable(symbol_ok = FALSE) ~  # condition : selection is code
    call_addin("reprex", "Reprex selection"),   # action    : call existing addin
)
```

We see below that whenever the selected code is parsable and is more than just a symbol,
the action will be proposed.

*INSERT GIF*

`selection_is_parsable()` is another **condition helper**. `call_addin()` is an
**action helper**, there are more of them documented in `?``action-helpers```.

It is good practice to have narrow conditions around our use cases 
so as our list of tricks grows we only display those that apply.
We could have defined `tricks::add_tricks("Reprex selection" = TRUE ~ call_addin("reprex", "Reprex selection"))` but then Reprex Selection would always be proposed even when it doesn't apply.

## `debugonce()`

You might not like to type the name of a function when you want
to debug, we might could define an addin to call `debugonce()` on your selection
but then you have to find an unused hotkey and remember it.

We can set a trick to `debugonce()` a function, it should be proposed only if
the selection is a symbol bound to a function.

```{r, eval = FALSE}
tricks::add_tricks(
  # label : here using glue syntax to have dynamic action names
  "debugonce({current_selection()})" = 
    # condition : selection evaluates to function
    selection_is_function() ~ 
    # action    : run debugonce, we use `bquote()`'s `.()` notation to work around NSE issues
    debugonce(.(current_call())),      
)

```

We see below that the action is proposed only if a symbol is selected and
evaluates to a function.

**INSERT GIF**

`selection_is_function()` is another **condition helper**.
`current_selection()` and `current_call()` are **context informers**, they are documented
in `?current_selection`.

We note a couple new things here :
* We used {glue} notation in the label, this is handy to provide custom labels
* We used `.()` inside `debugonce()`, this is the same `.()` that we find in 
`bquote()`, `debugonce()` uses NSE and supporting `.()` in actions and conditions
  permit us to program with such functions conveniently.

## Where are these tricks stored ? Can I remove some ?

They're stored in the `"tricks.tricks"` option which you can display with `getOption("tricks.tricks")`
or `tricks::show_tricks()`. You can remove tricks by calling `tricks::rm_tricks()` with
the labels of the tricks to remove.

## Isn't it slow to have a lot of tricks ? Is it safe ?

All of the helper functions are memoised so they will only be called once, and results will
stay in memory while other conditions are evaluated. Given that
most of these operations are pretty quick to begin with it's unlikely that you'll witness
big lags. However if you push it you might succeed to design conditions that do take
non trivial time to evaluate.

We discourage evaluating the selection (unless it's a symbol) by forbidding the use
of `current_value()` in conditions. Evaluating a complex expression might take time and/or have side effects
and this would really spoil the fun of {tricks}. Unless you try really hard, conditions
won't affect the global state so {tricks} can be used safely.

Moreover condition cannot trigger an error, if its code fails it just returns `FALSE`, 
the evaluation of conditions is also totally silent.



## Use packages of tricks or design your own

`tricks::add_tricks()` can take named formulas as arguments as we've seen above.
It can also take `"trick"` objects.

`"trick"` objects are created using the `new_trick()` function and they can be fed as
unnamed arguments.

If we want to package the first trick we defined in this document we might create
a package {mytricks} and have : 

```{r, eval = FALSE}
#' Edit user R Profile
#' 
#' A trick that proposes edit the user R profile
#'@export
#'@name edit_user_rprofile
edit_user_rprofile <- tricks::new_trick(
  "Edit user '.Rprofile'",
  ~ selection_is_empty(),
  ~ usethis::edit_r_profile()
)
```

And then tricks can be added this way 
```{r, eval = FALSE}
tricks::add_trick(
  my.trick.pkg::edit_user_rprofile,
)
```
